---
title: "Chapter 1"
author: "Team 2"
date: "3/29/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
Chapter 1 demonstrates how to access the Longitudinal Tract Data Base (LTDB) data dictionary, which was wrangled from multiple decennial data files. Three functions demonstrate the ability to filter the large number of LTDB variables to select only those with a certain category title, where the description field contains a specified text string, or from a specific year.


```{r, message=FALSE, warning=FALSE }

# load needed packages
library( tidyverse )
library( here )
library( pander )


```



```{r}

# store data dictionary file path
DD_FILEPATH <- here::here( "data/rodeo/LTDB-DATA-DICTIONARY.csv" )

# import data dictionary
dd <- read.csv( DD_FILEPATH, stringsAsFactors=F )

dd <- select( dd, c( 1:14))

dd[ dd == "" ] <- NA

```

## Building and Testing Filters

<br>

### Filter by category

```{r}

# function filters by category

filter_cat <- function( data , search )
  
{
  num <- grep(search, data$category, ignore.case = T) 
  
  dat.cat <- data.frame(data[num,])
  
  return( dat.cat )
}

```

```{r}

# test function 1, filtering for category "age"

filter_cat( dd, "age" )

```
<br>
### Filber by text in description field


```{r}
# function 2 filters by description text

search_var <- function(string)
{
vector <- dd$definition

these <- grepl( string, vector, ignore.case=T )

dat.sub <- dd[ these, ]

dat.sub <- select( dat.sub, c( 5:14 ) )

dat.sub <- dat.sub[ !is.na( dat.sub )]

return( dat.sub )

}

```


```{r}
# test function 2, returns all variables with the word "race" in description text

search_var( "race")

```
<br>

### Filter by year


```{r}
# function 3 filters by year, returns variables available for that year

filter_year <- function( year )
{
  
  dd.year <- dd %>% dplyr::select( contains( year ) )
  
  dd.year <- dd.year[ !is.na( dd.year )]

  return( dd.year )
  
}

```


```{r}
# test function 3, filtering for year 1970

filter_year( "1970" )

```




